---
title: "hw02-gapminder"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gapminder)
library(tidyverse)
```

##What is Gapminder, anyway?
```{r}
class(gapminder)
```
Gapminder is a data frame, so it should have numerous variables arranged in unique columns that each have the same number of rows.
```{r}
summary(gapminder)
nrow(gapminder)
```

The **6** columns in the Gapminder data frame are

 - Country (country)
 - Continent (continent)
 - Year (year)
 - Life Expectancy (lifeExp)
 - Population (pop)
 - GDP per Capita (gdpPercap)
 
There are **1704** observations for each variable, and so 1704 rows.Each variable is a list of either categorical (country, continent) or quantitative (pop, lifeExp) values.

##Exploring Gapminder

####Categorical
Looking at the Continent column, we can see there are only a few unique values this variable can take:
```{r}
cont <- select(gapminder, continent)
unique(cont)
```
The histogram below provides a summary of the distribution of values throughout the "continent" column. 
```{r}
ggplot(gapminder, aes(continent)) +
  stat_count()
```

####Quantitative
GDP per Capita is fairly different from the Continent variable, as it can assume any number within a certain range. In the table below we can see the minimum value of GDP per capita is $241.20, the maximum is $113,523.10. The mean GDP per Capita is $7215.30, which is much closer to the minimum than the maximum (75% of the data is below $9325.5).

```{r}
select(gapminder, gdpPercap) %>% 
  summary()
```

The density plot below visualizes the distribution of GDP per Capita.
```{r}
ggplot(gapminder, aes(gdpPercap)) +
  geom_density()
```

####Filter
Would how would the plot change if we only look at observations from Europe?
```{r}
gapminder %>% 
  filter(continent == "Europe") %>% 
  ggplot(aes(gdpPercap)) +
  geom_density()
```



####What might the GDP per Capita look like for different continents?

```{r}
ggplot(gapminder, aes(gdpPercap)) +
  geom_histogram() +
  facet_wrap( ~ continent)
```

